Interisland movements of woodland caribou in Newfoundland

### 2020-02-21 paper/manuscript.Rmd
Update manuscript.Rmd

The Fogo Island archipelago, off the coast of Newfoundland, Canada (49°N, 54°W),
is home to approximately 300 woodland caribou. During routine fieldwork on May
30, 2017, we observed an unmarked adult male caribou swim between Perry and
Eastern Indian Islands (Figure 1), a swim of at least 470 m which took ~9
minutes (~52 m per minute). This observation prompted us to investigate whether
swimming behaviour was common among GPS radio-collared caribou in the population
(n = 29 adult females: 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket,
Ontario, Canada). We identified swimming events when two consecutive GPS (2-hour
relocation rates) locations from an individual occurred on different islands.
This region typically experiences pack ice during the winter, so caribou may be
able to travel between islands by walking over the ice. We restricted our GPS
data to the ice-free period of the year (April 1 - November 25) to avoid this
possibility. Using these criteria, we identified three adult female caribou that
collectively engaged in 86 swimming events (Figure 2).

<!--Is it better to just state our results here, or give this sort of teaser? Initially I had "86 events by 3 individuals" here, but then we ignore that for a couple paragraphs of context before getting back into it... which seemed weird-->

Caribou were introduced to Fogo Island (n = 26 animals between 1964–67) and
Change Island (n = 5 animals in 1964) from the main Island of Newfoundland as
part of a series of translocations and introductions throughout the province
(Bergerud & Mercer 1989). Since 2016 we have been studying the Fogo Island
caribou herd (Bonar et al. 2017; Bonar et al. 2018; Peignier et al. 2019). To
date, we have fitted 29 adult female caribou with GPS collars (1240g, GPS 4400M;
Lotek Wireless Inc., Newmarket, Ontario, Canada), which were programmed to
collect relocation fixes every 2 hours (see Bonar et al. 2018 for details). We
used GPS locations from these individuals to investigate oceanic swimming
events.

>>>>

The Fogo Island archipelago, off the coast of Newfoundland, Canada (49°N, 54°W),
is home to approximately 300 woodland caribou. During routine fieldwork on May
30, 2017, we observed an unmarked adult male caribou swim between Perry and
Eastern Indian Islands (Figure 1), a swim of at least 470 m which took ~9
minutes (~52 m per minute). This observation prompted us to investigate whether
swimming behaviour was common among GPS radio-collared caribou in the population
(n = 29 adult females: 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket,
Ontario, Canada). We identified swimming events when two consecutive GPS (2-hour
relocation rates) locations from an individual occurred on different islands.
This region typically experiences pack ice during the winter, so caribou may be
able to travel between islands by walking over the ice. We restricted our GPS
data to the ice-free period of the year (April 1 - November 25) and we
identified three adult female caribou that collectively engaged in 86 swimming
events (Figure 2).

In total, 
<!--Is it better to just state our results here, or give this sort of teaser? Initially I had "86 events by 3 individuals" here, but then we ignore that for a couple paragraphs of context before getting back into it... which seemed weird-->

<!-- Caribou were introduced to Fogo Island (n = 26 animals between 1964–67) and
Change Island (n = 5 animals in 1964) from the main Island of Newfoundland as
part of a series of translocations and introductions throughout the province
(Bergerud & Mercer 1989). Since 2016 we have been studying the Fogo Island
caribou herd (Bonar et al. 2017; Bonar et al. 2018; Peignier et al. 2019). To
date, we have fitted 29 adult female caribou with GPS collars (1240g, GPS 4400M;
Lotek Wireless Inc., Newmarket, Ontario, Canada), which were programmed to
collect relocation fixes every 2 hours (see Bonar et al. 2018 for details). We
used GPS locations from these individuals to investigate oceanic swimming
events. -->
